Solutions to the Assignment - 2 : CS6230 - \\ Optimization Methods in Machine Learning

### 2017-10-10 Assignment-2/Solns-Assgn-2.tex
Assgn-2 Update 4

\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{bookmark}
\usepackage{hyperref}

@@

@@

\subsection*{Part c}
\begin{flushleft}
We know that if \(z = \textbf{prox}_{h, t}(x)\), then:
\begin{equation}
\label{subgrad-optim}
x - z \in \partial h(z) \implies x - \textbf{prox}_{h, t}(x) \in \partial h(\textbf{prox}_{h, t}(x))
\end{equation}

Substitute \(x \leftarrow x - t\nabla g(x)\) in \ref{subgrad-optim} to get:
\begin{equation}
\label{update-subgrad-optim}
x - t\nabla g(x) - \textbf{prox}_{h, t}(x - t\nabla g(x)) \in \partial h(\textbf{prox}_{h, t}(x - t\nabla g(x)))
\end{equation}

Using the definition of \(\textbf{prox}_{h, t}(x - t\nabla g(x)) = x - tG_{t}(x)\) in \ref{update-subgrad-optim}, we get:
\begin{equation}
\label{sub-part-c}
x - t\nabla g(x) - x + tG_{t}(x) \in \partial h(x - tG_{t}(x)) \implies tG_{t}(x) - t\nabla g(x) \in \partial h(x - tG_{t}(x)) 
\end{equation}

>>>>

\documentclass[11pt]{article}
\usepackage[margin=0.8in]{geometry}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{bookmark}
\usepackage{hyperref}

@@

\section*{Question 3}
\subsection*{Part a}
\begin{flushleft}
Consider the SVD of \(W\) to be \(W = U_{W} \Sigma_{W} V^{T}_{W}\). From the conditions imposed on \(W\), we get:

Hence the SVD of \(UV^{T} + W\) can be given by:
\begin{equation}
\label{svd-uv-w}
UV^{T} + W = U I V^{T} + U_{W} \Sigma_{W} V^{T}_{W} = \begin{bmatrix} U & U_{W} \end{bmatrix} \begin{bmatrix} I & 0 \\ 0 & \Sigma_{W} \end{bmatrix} \begin{bmatrix} V^{T} \\ V_{W}^{T} \end{bmatrix}
\end{equation}

This is a valid SVD because:
\begin{equation*}
U^{T}W = 0 \implies U^{T}U_{W}\Sigma_{W}V^{T}_{W} = 0 \implies U^{T}U_{W}\Sigma_{W}V^{T}_{W}V_{W} = 0 \implies U^{T}U_{W}\Sigma_{W} = 0 \implies U^{T}U_{W} = 0
\end{equation*}

Similarly it can be shown that \(VV_{W}^{T} = 0\) (by using \(WU = 0\) and pre-multiplying \(U^{T}_{W}\)). From these results, it can be shown that:
\begin{gather*}
\begin{bmatrix} U & U_{W} \end{bmatrix}^{T} \begin{bmatrix} U & U_{W} \end{bmatrix} = I_{2r} \\
\begin{bmatrix} V & V_{W} \end{bmatrix}^{T} \begin{bmatrix} V & V_{W} \end{bmatrix} = I_{2r}
\end{gather*}

Since \(||W||_{\text{op}} \leq 1\), we can tell that the diagonal entries of \(\Sigma_{W}\) are \(\leq 1\). Hence the maximum diagonal element in the singular matrix for \(UV^{T} + W\) from \ref{svd-uv-w} is \(1\), which suggests that \(||UV^{T} + W||_{\text{op}} \leq 1\)
\end{flushleft}

\subsection*{Part b}
\begin{flushleft}
Using the SVD of \(A\), we have:
\begin{equation}
\label{uv-w-a-prod}
(UV^{T} + W)^{T}A = (VU^{T} + W^{T})(U\Sigma V^{T}) = VU^{T}U\Sigma V^{T} + W^{T}U\Sigma V^{T} = V\Sigma V^{T}
\end{equation}

Now, using the properties of circularity of the trace and \ref{uv-w-a-prod}, we get:
\begin{equation}
\label{part-3b}
\text{tr}((UV^{T} + W)^{T}A) = \text{tr}(V\Sigma V^{T}) = \text{tr}(V^{T}V\Sigma) = \text{tr}(\Sigma) = ||A||_{*}
\end{equation}
\end{flushleft}

\subsection*{Part c}
\begin{flushleft}
We know that the dual norm of the operator norm is the trace norm. Hence by the definition of trace norm:
\begin{equation}
\label{trace-norm-dual}
\displaystyle ||A||_{*} = \max_{||B||_{\text{op}} \leq 1} \text{tr}(B^{T} A)
\end{equation}
From the result in \ref{ans-1a}, \ref{trace-norm-dual} and \(\nabla_{A} \text{tr}(B^{T}A) = B\), we know how to obtain the subdifferential of \(||A||_{*}\). This happens to be:
\begin{equation}
\label{subdiff-trace}
\displaystyle \partial ||A||_{*} = \text{Conv}\left(\{B \ni \text{tr}(B^{T}A) = ||A||_{*}, ||B||_{\text{op}} \leq 1\}\right)
\end{equation}
From the results in \ref{part-3b} and \ref{svd-uv-w}, we can see that \(UV^{T} + W\) for any \(W\) satisfying the same condition as the elements of \(\partial ||A||_{*}\). Hence we can conclude that:
\begin{multline}
\label{subset-rel}
\text{Conv}\left(\{B \ni \text{tr}(B^{T}A) = ||A||_{*}, ||B||_{\text{op}} \leq 1\}\right) \\ \supseteq \{UV^{T} + W \ni ||W||_{\text{op}} \leq 1, U^{T}W = 0, WV = 0, \text{tr}((UV^{T} + W)^{T}A) = ||A||_{*}\}
\end{multline}

Using \ref{subset-rel} and \ref{subdiff-trace}, we get the final result:
\begin{equation}
\label{ans-3c}
\displaystyle \partial ||A||_{*} \supseteq \{UV^{T} + W \ni ||W||_{\text{op}} \leq 1, U^{T}W = 0, WV = 0, \text{tr}((UV^{T} + W)^{T}A) = ||A||_{*} \}
\end{equation}

\end{flushleft}

@@

\subsection*{Part c}
\begin{flushleft}
By subgradient optimality condition:
\begin{equation}
\label{subgrad-optim}
0 \in \partial\left(\frac{||z-x||^2}{2t} + h(z)\right)  \implies \frac{x - z}{t} \in \partial h(z) \implies \frac{x - \textbf{prox}_{h, t}(x)}{t} \in \partial h(\textbf{prox}_{h, t}(x))
\end{equation}

Substitute \(x \leftarrow x - t\nabla g(x)\) in \ref{subgrad-optim} to get:
\begin{equation}
\label{update-subgrad-optim}
\frac{x - t\nabla g(x) - \textbf{prox}_{h, t}(x - t\nabla g(x))}{t} \in \partial h(\textbf{prox}_{h, t}(x - t\nabla g(x)))
\end{equation}

Using the definition of \(\textbf{prox}_{h, t}(x - t\nabla g(x)) = x - tG_{t}(x)\) in \ref{update-subgrad-optim}, we get:
\begin{equation}
\label{sub-part-c}
\frac{x - t\nabla g(x) - x + tG_{t}(x)}{t} \in \partial h(x - tG_{t}(x)) \implies \frac{tG_{t}(x) - t\nabla g(x)}{t} \in \partial h(x - tG_{t}(x)) 
\end{equation}
